Formalization and Automation of Variants of G\"{o}del's Ontological Proof of God's Existence

### 2013-08-17 Abstract/GodProof-Abstract.tex
Add bibliographical reference for Coq
Add citations and compare axioms and definitions with Gödel's originals

\usepackage{latexsym}
\usepackage{bussproofs}
\EnableBpAbbreviations
\newcommand{\rl}[1]{\RightLabel{#1}}

@@

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  G\"{o}del's proof \cite{Goedel} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
%
G\"{o}del defines God as a being who possesses all \emph{positive} properties.
He does not extensively discuss what positive properties are, 
but instead he states a few reasonable but debatable axioms that they should satisfy.
Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof. Our formalization employs the following axioms\footnote{Say where they come from and how they relate to G\"odel.}: (1) Any property necessarily implied by a positive property is positive;
(2) A property is positive if and only if its negation is not positive; (3) the property of being God-like is positive; (4) positive properties are necessarily positive; and (5) necessary existence is a positive property.
Moreover, the following definitions are employed:
(D1) x is God-like if and only if x has every positive property, (D2) a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, (D3) x necessarily exists if and only if every essence of x is necessarily exemplified.

From these axioms and definitions we then infer:

\begin{description}
\item {(L1)} Positive properties are possibly exemplified.
\item {(L2)} Possibly God exists.
\item {(L3)} If x is God-like, then the property of being 
God-like is an essence of x.
\item {(T)} \ Necessarily God exists.
\end{description}

@@

The above variant of G\"{o}del's proof has now been 
constructed for the first-time
with the utmost degree of detail and formality; cf.~\cite{ToDo:GitHubRepository}. The following has been done:
\begin{itemize}
\item A detailed natural deduction proof.
%
\item A formalization of the axioms and theorems in the TPTP THF format \cite{J22}.
%
\item Automatic verification of the consistency of the axioms and 
definitions with Nitpick \cite{Nitpick}.
%
\item Automatic demonstration of the theorems with the provers LEO-II \cite{LEO-II} and Satallax \cite{Satallax}.

\item A step-by-step formalization using the Coq proof assistant \cite{ToDo}.

>>>>

\usepackage{url}

@@

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  G\"{o}del's proof \cite{Goedel1970} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
%
G\"{o}del defines God as a being who possesses all \emph{positive} properties.
He does not extensively discuss what positive properties are, 
but instead he states a few reasonable but debatable axioms that they should satisfy.
Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof (e.g. \cite{Scott,Sobel,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). Our formalization employs the following axioms and definitions\footnote{
(A1), (A2), (A5), (D1), (D3) are logically equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and 3 in G\"odel's manuscript \cite{Goedel1970}. (A3) was introduced by Scott \cite{Scott} and could be derived from G\"odel's axiom 1 and (D1) in a logic with infinitary conjunction. (A4) is a weaker form of G\"odel's axiom 3. (D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
}: 
ToDo: Include formulas too??
\begin{description}
\item {(A1)} Any property necessarily implied by a positive property is positive;
\item {(A2)} A property is positive if and only if its negation is not positive; 
\item {(A3)} the property of being God-like is positive; 
\item {(A4)} positive properties are necessarily positive; 
\item {(A5)} necessary existence is a positive property.
\item {(D1)} x is God-like if and only if x has every positive property, 
\item {(D2)} a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, 
\item {(D3)} x necessarily exists if and only if every essence of x is necessarily exemplified.
\end{description}

\noindent
From these axioms and definitions we then infer:
%
ToDo: Include formulas too??
\begin{description}
\item {(L1)} Positive properties are possibly exemplified.
\item {(L2)} Possibly God exists.
\item {(L3)} If x is God-like, then the property of being 
God-like is an essence of x.
\item {(T)} \ Necessarily God exists.
\end{description}

@@

The above variant of G\"{o}del's proof has now been 
constructed for the first-time
with an unprecedent degree of detail and formality; cf.~\cite{FormalTheologyRepository}. The following has been done:
\begin{itemize}
\item A detailed natural deduction proof.
%
\item A formalization of the axioms and theorems in the TPTP THF format \cite{J22}.
%
\item Automatic verification of the consistency of the axioms and 
definitions with Nitpick \cite{Nitpick}.
%
\item Automatic demonstration of the theorems with the provers LEO-II \cite{LEO-II} and Satallax \cite{Satallax}.

\item A step-by-step formalization using the Coq proof assistant \cite{Coq}.
